An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-22 trunk/DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
minor improvements

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% our packages and definitions: %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{url} 
\usepackage{color}
% \usepackage{dsfont}

@@

\begin{figure}%
\includegraphics[width=\columnwidth]{figs/truckpaths}%
\caption{\label{fig:paths} Illustration of the variable-length representation of a solution (left) with its two paths for commodity 1 and one path for commodity 2 in opposition to the resulting full source--destination paths where the last part is the shortest path between the variable-length path's end node and the destination node for this commodity (right). The source and destination nodes for commodity 1 are 0 and 6 and for the second commodity the source is node 1 and the destination is node 7. The smaller numbers along the arcs depict the costs.}
\end{figure}

@@

\subsubsection{A Simple Random Initialization}
Based on the idea of representing a solution by the beginning of the truck paths which are completed by the shortest path to the destination nodes, we propose the following random initialization:
\begin{enumerate}
	\item Choose for each truck a node in the network uniformly at random among all nodes which are not the source.
	\item Initialize the truck paths with the shortest path from the associated source node to the randomly selected node.
\end{enumerate}
Note that this initialization scheme is not fully random as the initialization results in truck paths which are constructed by concatenating two shortest paths---the shortest path from the source node to the one randomly selected and the shortest path from the random node to the destination, which is added due to the representation described above.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following experiments have two goals:
- show that the algorithm works in principle on practical examples in manageable time
- check which of the two initialization schemes is 
In addition, a few more experiments show that the standard parameters of the algorithm behaves as with other test problems: population size, environmental selection (change from HypE to NSGA-II)

As a test case, we chose to use a real-world case study based on the work of \citet{bianco09}. The network used here consists of 311 nodes and 441 arcs and represents the main transport roads in Italy's Lazio region which amongst other cities includes the capital Rome.

The risks $r_{ij}^c$ for each arc $(i,j) \in A$ and commodity $c \in C$ are evaluated as the societal risk computed as the number of people living inside the exposure zone around link $(i,j)$ (whose size depends on the hazmat type $c$ times the accident probability involving the vehicle), for details see  \citep{bianco09}.

@@

@@

\subsection{The Feasibility of the EMO Approach}
\begin{itemize}
	\item of course no guarantee that the solutions are Pareto-optimal
	\item however, in reasonable time (e.g. ??hrs for 10 runs with 1000 generations each) results with several non-dominated solutions the decision maker can interpret and with which she can get an idea about the trade-offs between the objectives
	\item if DM wants good results fast, cost-efficient initialization is favourable, otherwise, random initialization gives better spread of the solutions
	\item if needed, approach can be combined easily with preference articulation techniques such as the weighted hypervolume indicator which allows to steer the search towards user-defined regions of the objective space (show 1 plot with W-HypE).
	\item approach is very general and can easily deal with slightly modified problems: e.g. different truck loads (only the objective function code needs to be changed)
\end{itemize}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% our packages and definitions: %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{url} 
\usepackage{color}
% \usepackage{dsfont}
\usepackage{booktabs}
\usepackage{amsmath}

@@

\begin{figure}%
\includegraphics[width=\columnwidth]{figs/truckpaths}%
\caption{\label{fig:paths} Illustration of the variable-length representation of a solution (left) with its two paths for commodity 1 and one path for commodity 2 in opposition to the resulting full source--destination paths where the last part is the shortest path between the variable-length path's end node and the destination node for this commodity (right). The source and destination nodes for commodity 1 are 0 and 6 and for the second commodity the source is node 1 and the destination is node 7. The smaller numbers along the arcs of the graph in the middle depict the costs of the instance.}
\end{figure}

@@

\subsubsection{A Simple Random Initialization}
Based on the idea of representing a solution by the beginning of the truck paths which are completed by the shortest path to the destination nodes, we propose the following random initialization:
\begin{enumerate}
	\item Choose for each truck a node in the network uniformly at random among all nodes which are not the source.
	\item Initialize the truck paths with the shortest path from the associated source node to the randomly selected node.
\end{enumerate}
Note that this initialization scheme is, though random, does not produce fully random paths. The initialization results in truck paths which are constructed by concatenating two shortest paths---the shortest path from the source node to the one randomly selected and the shortest path from the random node to the destination, which is added due to the representation described above.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the remainder of this paper, we investigate how the above proposed evolutionary multiobjective optimization algorithm can cope with realistic instances of the multiobjective risk-equity constrained routing problem. The experimental analyses have two main goals. On the one hand, we show that the algorithm is efficient and can find multiple non-dominated solutions on practical examples in reasonable time. On the other hand, we investigate which of the two proposed initialization schemes is favourable and when this is the case. In addition, a few more experiments are conducted to show that the standard parameters of the algorithm behave as with other test problems. To this end, we vary the population size of the algorithm as well as its environmental selection scheme.

As a test case for all the experiments, we chose to use a real-world case study based on the work of \citet{bianco09}. The network used here consists of 311 nodes and 441 arcs and represents the main transport roads in Italy's Lazio region which, amongst other cities, includes the capital Rome. We distinguish between instances with a different number of commodities (between 2 and 4) and with different numbers of trucks, see Table~\ref{tab:instances}. While the costs $c_{ij}$ are proportional to the real distance between the two locations $i$ and $j$, the risks $r_{ij}^c$ for each arc $(i,j) \in A$ and commodity $c \in C$ have been computed as a product of the arc's length, the actual population density in the area of the arc and a constant which depends on the commodity, for details see \citep{bianco09} and Table~\ref{tab:instances}. Solely one instance is created differently: to show the impact of correlation between the costs and the risks, we created the instance lazio$_{\text ncorr}$\_4\_1 with the same graph and risks as for the instance lazio\_4\_1 but with randomly permutated costs where each occuring cost value has been divided by a random value between 0 and 5000.

\begin{table}
\caption{\label{tab:instances} Parameters of the instances used in the experimental analyses.}
\begin{tabular}{lllll}
	\toprule
	name & \begin{minipage}{1cm}\#com\-mod\-i\-ties\end{minipage} & \#trucks & cost range & risk range \\ \midrule
	lazio\_2\_1 & 2 & \small 20+19         & & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\end{minipage}\\
	lazio\_2\_2 & 2 & \small 23+29         & & \\
	lazio\_3\_1 & 3 & \small 25+16+28      & & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\end{minipage}\\
	lazio\_4\_1 & 4 & \small 21+3+30+5     & \small [0.25, 41] & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\\ x [762, 274628]\end{minipage}\\
	lazio$_{\text uncorr}$\_4\_1 & 4 & \small 21+3+30+5     & \small [0.0109239, 3.93701] & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\\ x [762, 274628]\end{minipage}\\
	\bottomrule
\end{tabular}
\end{table}

@@

\subsection{Experimental Settings}
Except otherwise stated, we used the following parameter setting. We perform 10 independent runs of HypE for 2000 generations with a population size of 100. The reference point for the hypervolume indicator was chosen as in Table~\ref{tab:referencepoints} according to the initial objective function values, such that all initial solutions have a positive contribution to the hypervolume indicator\footnote{Note that the reason for using the same values in all objectives is the used PISA implementation of HypE.}. In all cases, the hypervolume computation is exact.

\begin{table}
\centering
\caption{\label{tab:referencepoints} Used reference points within the environmental selection of HypE}
\begin{tabular}{lr}
	\toprule
	instance & reference point\\ \midrule
	\# commodities: 2 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 3 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 4 & $(10^9, 10^9, 10^9)$\\
	\bottomrule
\end{tabular}
\end{table}

@@

\subsection{Changing the Population Size}
In a second experiment, we show that changing the population size of the algorithm in standard ranges does not change the algorithm outcome drastically.
\TODO{write}

\subsection{Changing the Environmental Selection}
With a last experiment, we show the influence of the algorithm's environmental selection scheme on the performance and in particular how this can be used to steer the search towards specific regions of the Pareto front.
\TODO{write}

\subsection{The Feasibility of the EMO Approach}
\begin{itemize}
	\item of course no guarantee that the solutions are Pareto-optimal
	\item however, in reasonable time (e.g. ??hrs for 10 runs with 1000 generations each) results with several non-dominated solutions the decision maker can interpret and with which she can get an idea about the trade-offs between the objectives
	\item if DM wants good results fast, cost-efficient initialization is favourable, otherwise, random initialization gives better spread of the solutions
	\item if needed, approach can be combined easily with preference articulation techniques such as the weighted hypervolume indicator which allows to steer the search towards user-defined regions of the objective space (show 1 plot with W-HypE).
	\item approach is very general and can easily deal with slightly modified problems: e.g. different truck loads (only the objective function code needs to be changed)
	\item it needs to be mentioned that for the tested instances with 311 nodes and 441 arcs, the variation operator of the evolutionary algorithm was able to construct the shortest paths from source to destination in some of the runs for some of the truck paths within the first 100,000 function evaluations. This corresponds to performing random walks of lengths around 20 nodes.
\end{itemize}
